Model Finding for Recursive Functions in SMT\thanks{ This research is partially supported by the Inria technological development action ``Contre-exemples Utilisables par Isabelle et Coq'' (CUIC). }

### 2015-05-07 tex/main.tex
more

First-order logic has built-in notion of computation or termination. To ensure
that a function specification is well-founded, it is sufficient to require that
it would terminate when seen as a functional program for \relax{some}
evaluation order. For example, the definition~(\ref{eq:pii}) from the
introduction can be shown well-founded under an evaluation strategy that
consides one of the two $\lite$ branches, based on the result of evaluating the
condition. Logically, such dependencies can be captured by congruence rules.
Krauss develops these ideas in the more general context of
higher-order logic \cite[Section 2]{krauss-2009-phd}.

\begin{theorem}\label{thm:adm}
If
$\DDD$ is a set of well-founded function definitions for
$\sfundefs{\Sigma} = \{\con{f}_1,\ldots,\con{f}_n\}$, then it is admissible.
\end{theorem}
\begin{proofsketch}
Let $\psi$ be a satisfiable formula closed under function expansion with
respect to $\DDD$. We show that $\psi \wedge \bigwedge \Delta$ is also
satisfiable. Let $\I$ be a model of $\psi$, and
let $\J_0$ be the restriction of $\I$ to the function symbols in $\sfuns{\Sigma} -
\sfundefs{\Sigma}$. Because well-founded definitions uniquely characterize
the interpretation of the functions they define, there exists a
$\Sigma$-interpretation $\J$ such that $\J \models \Delta$.
%
Since $\psi$ is closed under function expansion, it already constraints the
functions in $\sfundefs{\Sigma}$ recursively as far as is
necessary for interpreting $\psi$. Thus, any point $v$ for which
$\const{f}_i^\I(v)$ is queried when interpreting $\psi$ will have its
expected value according to its definition and hence coincide with $\J$.
And since $\psi^\I$ does not depend on the interpretation at the other
points, $\J$ is, like $\I$, a model of $\psi$. Furthermore, $\J \models \Delta$
by assumption. Thus, $\J \models \psi \wedge \bigwedge \Delta$.
%We have
%$\varphi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \varphi ) \}$.
\qed
\end{proofsketch}

>>>>

First-order logic has no built-in notion of computation or termination. To ensure
that a function specification is well-founded, it is sufficient to require that
the function would terminate when seen as a functional program, under \emph{some}
evaluation order. For example, definition~(\ref{eq:pii}) from the
introduction can be shown well-founded under an evaluation strategy that
consides one of the two $\lite$ branches, based on the result of evaluating the
condition. Logically, such dependencies can be captured by congruence rules.
Krauss developed these ideas in the more general context of
higher-order logic \cite[Section 2]{krauss-2009-phd}.

\begin{theorem}\label{thm:adm}
If\/
$\DDD$ is a set of well-founded function definitions for\/
$\sfundefs{\Sigma} = \{\con{f}_1,\ldots,\con{f}_n\}$, then it is admissible.
\end{theorem}
\begin{proofsketch}
Let $\psi$ be a satisfiable formula closed under function expansion with
respect to $\DDD$. We show that $\psi \wedge \bigwedge \Delta$ is also
satisfiable. Let $\I$ be a model of $\psi$, and
let $\J_0$ be the restriction of $\I$ to the function symbols in $\sfuns{\Sigma} -
\sfundefs{\Sigma}$. Because well-founded definitions uniquely characterize
the interpretation of the functions they define, there exists a
$\Sigma$-interpretation $\J$ such that $\J \models \Delta$.
%
Since $\psi$ is closed under function expansion, it already constraints the
functions in $\sfundefs{\Sigma}$ recursively as far as is
necessary for interpreting $\psi$. Thus, any point $v$ for which
$\const{f}_i^\I(v)$ is queried when interpreting $\psi$ will have its
expected value according to its definition and hence coincide with $\J$.
And since $\psi^\I$ does not depend on the interpretation at the other
points, $\J$ is, like $\I$, a model of $\psi$.
Since $\J \models \Delta$, we have $\J \models \psi \wedge \bigwedge \Delta$,
as desired.
%We have
%$\varphi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \varphi ) \}$.
\qed
\end{proofsketch}
